### 2022-12-01 intro.Rnw
Lisa feedback

\begin{wrapfigure}{r}{0.34\textwidth}
  \centering
  \includegraphics[height = 8.9cm]{images/xkcd.png}
  \caption*{Replication studies as illustrated by Randall Munroe
    (\url{https://xkcd.com/242/}).}
\end{wrapfigure}

@@

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. % Effect
% estimates, standard error, confidence intervals, and $p$-values
All of these are routinely reported in researcher articles, and if one of them
is missing they typically can, under some assumptions, be calculated from the
others. Using summary statistics is also often the only possible way for
conducting the replication success analysis as the raw data from the original
study may not be available to the replicators. The most important statistic is
the effect estimate $\that$. It provides an estimate of the underlying effect
size $\theta$ which quantifies the true effect or association of an
intervention/exposure with an outcome variable. Typical effect sizes are mean
differences and correlations (for continuous outcomes), odds ratios (for binary
outcomes), or hazard ratios (for time to event outcomes). Depending on the type
of effect size, a transformation might be required so that the assumption of
approximately normally distributed effect estimates around the unknown effect
size $\theta$ (for large enough sample sizes) is justified. This could be, for
instance, the Fisher $z$-transformation for correlations or the
log-transformation for odds/hazard ratios \citep[chapter 11]{Cooper2019}. The
associated standard error $\sigma$ represents the statistical uncertainty of the
estimate. Under the assumption of (asymptotic) normality, confidence intervals
for $\theta$ and $p$-values for testing the null hypothesis
$H_{0} \colon \theta = 0$ can be computed as shown in Table~\ref{tab:meta}.

@@

Table~\ref{tab:rsdef} lists commonly used criteria for replication success in
terms of the summary statistics from Table~\ref{tab:meta}. The most popular
criterion is listed first and defines replication success by simultaneous
statistical significance of original and replication study along with their
effect estimates showing the same direction. This approach is also called the
``two-trials rule'' \citep{Senn2008} in drug regulation or ``vote-counting'' in
meta-analysis \citep{Cooper2019}. The criterion can similarly be defined through
one-sided $p$-values, so that the original and replication effect estimate are
not required to show the same direction as this is taken into account by the
$p$-values. Some replication projects \citep{Opensc2015, Errington2021} also
defined replication success via simultaneous non-significance of original and
replication study ($p_{o} > \alpha$ and $p_{r} > \alpha$). However, with this
definition ``replication success'' can almost always be achieved by conducting
original and replication study with very few samples so that the $p$-values are
large. The approach is also logically questionable as it could be seen as an
instance of the ``absence of evidence fallacy'' \citep{Altman1995} meaning the
failure to find evidence against the null hypothesis is erroneously interpreted
as evidence for the null hypothesis. Meta-analytic extensions of the
significance approach define replication success by significance of a combined
effect estimate.\footnote{The meta-analytic approach can also be given a
  Bayesian interpretation: A flat prior distribution for the underlying effect
  size $\theta$ is updated by the data from original and replication study.
  Replication success defined by a meta-analytic $p$-value being smaller than
  $\alpha$ is then equivalent to replication success via a Bayesian posterior
  probability $\Pr(\theta > 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
  respectively, $\Pr(\theta < 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$
  depending on the direction of combined estimate.} Typically, the assumption of
a common underlying effect size is seen as reasonable so that fixed-effect
meta-analysis is used for pooling. Random-effects meta-analysis has primarily
been used if more than one replication study of the same original study are
conducted since replicators are often interested in understanding
between-replication heterogeneity \citep[\eg{} in][]{Klein2018}.

@@

\subsubsection{Example: Reproducibility Project Cancer Biology}
I will now illustrate the assessment of replicability on data from the
Reproducibility Project: Cancer Biology \citep{Errington2021}. This large-scale
project attempted to replicate 53 landmark studies from the field of cancer
biology but due to various difficulties (\eg{} missing information from the
original studies or problems in conducting the experiments) only 23 of the 53
studies could be repeated. \citet{Errington2021} report that these experiments
led to data on 158 quantitative effects. However, from the data which they
provide only 132 quantitative effects are provided with original and replication
standardized mean difference effect estimates along with standard errors, and
only they will be used in the subsequent analyses.

@@

upset(rpcbUpset, sets = c("Significance", "Meta analytic \nsignificance",
                          "Relative \neffect size", "Prediction \ninterval"),
      order.by = "freq", mainbar.y.label = "Successful replications (combined criteria)",
      sets.x.label = "Successful replications \n(single criterion)", #, empty.intersections = "on")
      queries = list(list(query = intersects,
                          params = c("Significance",
                                     "Meta analytic \nsignificance",
                                     "Relative \neffect size",
                                     "Prediction \ninterval"),
                          color = "#1b9e77", active = TRUE)))
@
\caption{UpSet plot for data on 132 effects from the Reproducibility Project:
  Cancer Biology \citep{Errington2021} for which effect estimates and standard
  errors are available on standardized mean difference scale. Shown are the
  number of replication successes according to the different criteria from
  Table~\ref{tab:rsdef} and their combinations. A threshold of
  $\alpha = \Sexpr{round(100*alpha, 2)}\%$ and a relative effect size threshold
  of $d_{\text{min}} = \Sexpr{dmin}$ are used.}
\label{fig:upset}
\end{figure}

@@

Taken together, this analysis demonstrates that conclusions based on commonly
used replication success criteria often differ. It is not always clear cut
whether or not a replication is successful. Reducing replicability to a single
criterion without mentioning these difficulties, as often done by the popular
press \citep[\eg{} ``more than half of the findings did not hold up when
retested'' in][]{Carey2015}, is a simplification of the matter.
% that may harm the credibility of science.

\begin{landscape}
\begin{table}
  \centering
  \caption{Subset of results from the Reproducibility Project: Cancer Biology
    \citep{Errington2021}. (P, X, E) denotes effect number E from experiment
    number X, from original paper number P. Shown are original $\that_{o}$,
    replication $\that_{r}$, and pooled effect estimate $\that_{m}$ with 95\%
    confidence intervals, variance ratio $c = \sigma^{2}_{o}/\sigma^{2}_{r}$,
    relative effect estimate $d = \that_{r}/\that_{o}$, original
    \textit{p}-value $p_{o}$, replication \textit{p}-value $p_{r}$,
    meta-analytic \textit{p}-value $p_{m}$, $Q$-test \textit{p}-value $p_{Q}$,
    and sceptical \textit{p}-value $p_{\mathrm{S}}$. A level $\alpha = 5\%$ is
    used as threshold for replication success via \textit{p}-values, and a
    threshold $d_{\text{min}} = 1$ for the relative effect estimate.}
  \label{tab:RPCB}
% \resizebox{1\textwidth}{!}{%
<< "RPCB-table2", results = "asis" >>=
## select a subset with one effect per paper
set.seed(7)
rpcbTable <- rpcb |>
    na.omit() |>
    ## filter(smdo < 80, smdr < 80) |>
    group_by(paper) |>
    sample_n(size = 1) |>
    ungroup() |>
    arrange(pr2) |>
    select(id,
           smdoFormat, ## so,
           smdrFormat, smdmFormat,
           c, dFormat, ## sr,
           po2Format, pr2Format, pm2Format, pQFormat, ps2Format
           )

@@

Bayesian inference is an approach to statistical inference where Bayes' theorem
is used to make probability statements about unknown parameters based on the
observed data. The central quantity for doing so is the distribution of the
parameters conditional on the data, called the \emph{posterior distribution}. It
can be obtained from Bayes' theorem
\begin{align*}
  f(\theta \given \text{data})
  = f(\theta) \times \frac{f(\text{data} \given \theta)}{f(\text{data})}
\end{align*}
meaning that the \emph{prior distribution} for the parameter $\theta$ with
probability density/mass function $f(\theta)$ is multiplied by the (normalized)
likelihood of the data, also know as \emph{Bayesian updating}. Parameter values
which increase the likelihood of the data become more likely \emph{a posteriori}
but they are also weighted by their \emph{a priori} plausibility through the
prior. As such, Bayesian inference provides a formal way for combining
information from the data at hand with external knowledge encoded in the prior.

Many consider the prior to be also the weak point of Bayesian inference since it
is unclear how it should be specified in the absence of external knowledge. The
\emph{reverse-Bayes} approach, first proposed by \citet{Good1950}, is one way of
dealing with this issue. The idea is to reverse Bayes' theorem
\begin{align*}
  f(\theta) = f(\theta \given \text{data}) \times
  \frac{f(\text{data})}{f(\text{data} \given \theta)}
\end{align*}
and instead ``downdate'' a posterior with the observed data. So, in contrast to
the conventional ``forward-Bayes'' approach where we start with a prior, update
it with the data, and end up with a posterior, the reverse-Bayes approach starts
with the posterior and ends up with the prior. For example, the posterior may be
specified to indicate evidence for/against a particular hypothesis under
investigation. Reverse-Bayes inference then revolves around the question of
whether the resulting prior is plausible in light of external knowledge, and if
so, this could be seen as support for the specified posterior.

To illustrate reverse-Bayes inference, let us return to the replication setting.
Assume we want to conduct inference about the unknown effect size $\theta$ based
on the effect estimate from the original study $\that_{o}$ and its standard
error $\sigma_{o}$. We will assume that $\that_{o}$ is normally distributed
around the unknown effect size $\theta$ with (known) variance equal to its
squared standard error $\sigma^{2}_{o}$, here and henceforth denoted by
$\that_{o} \given \theta \sim \Nor(\theta, \sigma^{2}_{o})$. Furthermore, we
specify a zero-mean normal prior with variance $\tau^{2}$ for the effect size
$\theta$, representing the position of a sceptic who does not believe in the
presence of a genuine effect. The ``stubbornness'' of the sceptic is determined
by how large the variance $\tau^{2}$ is chosen. Combining the sceptical prior
with the likelihood produces a posterior which is again normal
$\theta \given \that_{o}, \sigma_{o} \sim \Nor(\mu_{\text{post}}, \sigma^{2}_{\text{post}})$
with mean and variance
\begin{align*}
  &\mu_{\text{post}} = \frac{\that_{o}}{1 + \sigma^{2}_{o}/\tau^{2}}&
  &\text{and}&
  &\sigma^{2}_{\text{post}} = \frac{1}{1/\sigma^{2}_{o} + 1/\tau^{2}}.&
\end{align*}
The associated $(1- \alpha)$ highest posterior density credible interval is
given by
\begin{align}
  \left[\mu_{\text{post}} \pm z_{\alpha/2} \, \sigma_{\text{post}}\right]
  \label{eq:postCI}
\end{align}
and if this credible interval excludes parameter values smaller/larger than zero
(depending on the orientation of the effect size) this may be interpreted as
evidence\footnote{For readers who do not agree with this notion of evidence, do
  not panic. In this thesis (Paper II and IV) we extend this reverse-Bayes
  procedure to use alternative measures of evidence, such as Bayes factors.} for
a genuine effect found in the original study.

Depending on how large the prior variance $\tau^{2}$ is chosen, the posterior
credible interval~\eqref{eq:postCI} will either include or exclude zero.
Different researchers may have different degrees of scepticism and may thus
choose different prior variances $\tau^{2}$. As a default choice,
\citet{Held2020} proposed to use the reverse-Bayes approach from
\citet{Matthews2001b}, that is, to determine the \emph{sufficiently sceptical
  prior variance} $\tau^{2}_{\alpha}$ so that the appropriate limit of the
$(1 - \alpha)$ credible interval is just fixed to zero. The resulting prior then
represent the beliefs of a sceptic who is just stubborn enough to not find the
original study convincing at level $\alpha$.

@@

\citet{Held2019a} showed that the sufficiently sceptical prior variance
$\tau^{2}_{\alpha}$ for a level $\alpha$ is available in closed-form
\begin{align}
  \tau^{2}_{\alpha} =
  \begin{cases}
    \dfrac{\sigma^{2}_{o}}{\left(z_{o}^{2}/z_{\alpha/2}^{2}\right) - 1}
    & \text{if} ~ z_{o}^{2} > z_{\alpha/2}^{2} \\
    \text{undefined} & \text{else.}
  \end{cases}
  \label{eq:tau2alpha}
\end{align}
From~\eqref{eq:tau2alpha} we see that convincing original studies (those with
large absolute $z$-values $|z_{o}|$) require smaller sufficiently sceptical
prior variances to render the posterior no longer convincing for the same level
$\alpha$. Conversely, if the original study is not convincing enough (if
$|z_{o}| \leq z_{\alpha/2}$) the sufficiently sceptical prior variance is
undefined meaning that the data provide so little evidence that no scepticism is
required to make them unconvincing.

@@

Figure~\ref{fig:RBillustration} shows the prior predictive $p$-values
from~\eqref{eq:pboxNorm} computed for three example replication studies from the
Replication Project: Cancer Biology. We see that larger effect estimates show
smaller prior predictive $p$-values $p_{\text{Box}}$, indicating more prior-data
conflict. This is because the standard errors from all three replications are
roughly the same size, so that the distance between zero and the replication
effect estimate matters most. The $p$-values suggest that there is hardly any
conflict between the sceptical prior and the first two replications (those with
$p_{\text{Box}} = \Sexpr{pboxFormat[1]}$ and
$p_{\text{Box}} = \Sexpr{pboxFormat[2]}$), while the conflict seems larger for
the third one (the one with $p_{\text{Box}} = \Sexpr{pboxFormat[3]}$).

@@

zr <- tr/sr
c <- so^2/sr^2
ps <- pSceptical(zo = zo, zr = zr, c = c, alternative = "two.sided", type = "nominal")
psFormat <- pFormat(p = ps)
@
\subsubsection{The sceptical \textit{p}-value}
To remove the dependence on the level $\alpha$, \citet{Held2020} proposed to
determine the smallest level at which replication success can be established.
This level is called the \emph{sceptical $p$-value} $p_{\text{S}}$, and
available in closed-form
\begin{align*}
  p_{\text{S}}
  % = \inf_{\alpha \in (0, 1)} \left\{\alpha : p_{\text{Box}} \leq \alpha \right\}.
  = 2\left\{1 - \Phi(|z_{\text{S}}|)\right\}
\end{align*}
where
\begin{align*}
  z_{\text{S}}^{2}
  = \begin{cases}
    z^{2}_{\text{H}}/2 & \text{for} ~ c = 1 \\
    \left\{\surd\left[z_{\text{A}}^{2}\left\{z_{\text{A}}^{2} +
          z_{\text{H}}^{2}(c - 1)\right\}\right]
      - z_{\text{A}}^{2}\right\}/(c - 1)
    & \text{for} ~ c \neq 1
  \end{cases}
\end{align*}
with arithmetic mean $z_{\text{A}}^{2} = (z^{2}_{o} + z^{2}_{r})/2$ and harmonic
mean $z_{\text{H}}^{2} = 2/(1/z^{2}_{o} + 1/z^{2}_{r})$ of the squared
$z$-statistics, and variance ratio $c = \sigma^{2}_{o}/\sigma^{2}_{r}$.
Replication success at level $\alpha$ is then equivalent to
$p_{\text{S}} \leq \alpha$. For instance, the sceptical $p$-values of the three
replication studies in Figure~\ref{fig:RBillustration} are
$p_{\text{S}} = \Sexpr{psFormat[1]}$, $p_{\text{S}} = \Sexpr{psFormat[2]}$, and
$p_{\text{S}} = \Sexpr{psFormat[3]}$ (from left to right), so we can see that
the third replication is unsuccessful at level $\alpha = 5\%$ but successful at
level $\alpha = 10\%$. However, the sceptical $p$-value does not necessarily
have to be dichotomized in this manner, but can also be interpreted as a
quantitative measure of replication success, the smaller $p_{\text{S}}$ the
higher the degree of replication success.

The sceptical $p$-value has several interesting properties \citep[section
3]{Held2020}: First, it is always larger than the maximum of the original and
replication $p$-values ($p_{\text{S}} > \max\{p_{o}, p_{r}\}$), meaning that
both $p$-values have to be smaller than $\alpha$ such that replication success
at level $\alpha$ is possible. The sceptical $p$-value hence requires
\emph{both} studies to be sufficiently convincing on their own (in terms of
their $p$-values), similar to the significance criterion for replication
success. Second, if the $p$-values $p_{o}$ and $p_{r}$ remain fixed but the
relative effect estimate $d = \that_{r}/\that_{o}$ decreases, the sceptical
$p$-value increases, meaning that shrinkage of the replication effect estimate
is penalized ($p_{\text{S}} \uparrow 1$ as $d \downarrow 0$ for fixed $p_{o}$
and $p_{r}$). The sceptical $p$-value thus takes effect shrinkage into account,
unlike the significance criterion which can be achieved with any non-zero
replication effect estimate $\that_{r}$ provided the standard error $\sigma_{r}$
is small enough. This property is desirable in the replication setting as a
smaller replication effect estimate $\that_{r}$ may not be practically relevant
anymore, despite its statistical significance.

@@

This concludes the introduction to replication studies and the summary of the
reverse-Bayes assessment of replication success. Some additional properties
(\eg{} the null distribution of the sceptical $p$-value) and extensions (\eg{} a
one-sided version of the sceptical $p$-value, power and sample size
calculations) can be found in the original article by \citet{Held2020}.
Furthermore, the sceptical $p$-value and related methods are implemented in the
R package \texttt{ReplicationSuccess} available on CRAN
(\url{https://cran.r-project.org/package=ReplicationSuccess}). However, the
relevant parts of the procedure have been presented. In the following, I will
discuss open questions and how they are addressed in this thesis.

@@

\startPaperSummary{PaperI}{%
  The assessment of replication success based on relative effect size}{%
  Leonhard Held, Charlotte Micheloud, Samuel Pawel \\
  \textit{The Annals of Applied Statistics}, 2022, 16(2), 706--720.
  \url{https://doi.org/10.1214/21-AOAS1502}.}

>>>>

\begin{wrapfigure}{r}{0.34\textwidth}
  \centering
  \includegraphics[height = 9cm]{images/xkcd.png}
  \caption*{Replication studies as illustrated by Randall Munroe
    (\url{https://xkcd.com/242/}).}
\end{wrapfigure}

@@

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. % Effect
% estimates, standard error, confidence intervals, and $p$-values
All of these are routinely reported in research articles, and if one of them is
missing they typically can be calculated from the others. Using summary
statistics is also often the only possible way for conducting the replication
success analysis as the raw data from the original study may not be available to
the replicators. The most important statistic is the effect estimate $\that$. It
provides an estimate of the underlying effect size $\theta$ which quantifies the
true effect or association of an intervention/exposure with an outcome variable.
Typical effect sizes are mean differences and correlations (for continuous
outcomes), odds ratios (for binary outcomes), or hazard ratios (for time to
event outcomes). Depending on the type of effect size, a transformation might be
required so that the assumption of approximately normally distributed effect
estimates around the unknown effect size $\theta$ (for large enough sample
sizes) is justified. This could be, for instance, the Fisher $z$-transformation
for correlations or the log-transformation for odds/hazard ratios \citep[chapter
11]{Cooper2019}. The associated standard error $\sigma$ represents the
statistical uncertainty of the estimate. Under the assumption of (asymptotic)
normality, confidence intervals for $\theta$ and $p$-values for testing the null
hypothesis $H_{0} \colon \theta = 0$ can be computed as shown in
Table~\ref{tab:meta}.

@@

Table~\ref{tab:rsdef} lists commonly used criteria for replication success in
terms of the summary statistics from Table~\ref{tab:meta}. The most popular
criterion is listed first and defines replication success by simultaneous
statistical significance of original and replication study along with their
effect estimates showing the same direction. This approach is also called the
``two-trials rule'' \citep{Senn2008} in drug regulation or ``vote-counting'' in
meta-analysis \citep{Cooper2019}. The criterion can similarly be defined through
one-sided $p$-values, so that the original and replication effect estimate are
not required to show the same direction as this is taken into account by the
$p$-values. Some replication projects \citep{Opensc2015, Errington2021} also
defined replication success via simultaneous non-significance of original and
replication study ($p_{o} > \alpha$ and $p_{r} > \alpha$). However, with this
definition ``replication success'' can almost always be achieved by conducting
original and replication study with very few samples so that the $p$-values are
large. The approach is also logically questionable as it could be seen as an
instance of the ``absence of evidence fallacy'' \citep{Altman1995} meaning the
failure to find evidence against the null hypothesis is erroneously interpreted
as evidence for the null hypothesis. Meta-analytic extensions of the
significance approach define replication success by significance of a combined
effect estimate.\footnote{The meta-analytic approach can also be given a
  Bayesian interpretation: A flat prior distribution for the underlying effect
  size $\theta$ is updated by the data from original and replication study.
  Replication success defined by a meta-analytic $p$-value being smaller than
  $\alpha$ is then equivalent to replication success via a Bayesian posterior
  probability $\Pr(\theta > 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
  respectively, $\Pr(\theta < 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$
  depending on the direction of the combined estimate.} Typically, the
assumption of a common underlying effect size is seen as reasonable so that
fixed-effect meta-analysis is used for pooling. Random-effects meta-analysis has
primarily been used if more than one replication study of the same original
study are conducted since replicators are often interested in understanding
between-replication heterogeneity \citep[\eg{} in][]{Klein2018}.

@@

\subsubsection{Example ``Reproducibility Project: Cancer Biology''}
I will now illustrate the assessment of replicability on data from the
Reproducibility Project: Cancer Biology \citep{Errington2021}. This large-scale
project attempted to replicate 53 landmark studies from the field of cancer
biology \citep[and would perhaps therefore better be named ``Replication
Project'', see][]{Goodman2016}. In the end, only 23 of the 53 studies could be
repeated due to various difficulties, such as, missing information from the
original studies or problems in conducting the experiments.
\citet{Errington2021} report that these experiments led to data on 158
quantitative effects. However, from the data which they provide only 132
quantitative effects are provided with original and replication standardized
mean difference effect estimates along with standard errors, and only they will
be used in the subsequent analyses.

@@

upset(rpcbUpset, sets = c("Significance", "Meta analytic \nsignificance",
                          "Relative \neffect size", "Prediction \ninterval"),
      order.by = "freq", mainbar.y.label = "Successful replications (combined criteria)",
      sets.x.label = "Successful replications \n(single criterion)", #, empty.intersections = "on")
      queries = list(list(query = intersects,
                          params = c("Significance",
                                     "Meta analytic \nsignificance",
                                     "Relative \neffect size",
                                     "Prediction \ninterval"),
                          color = "#1b9e77", active = TRUE)))
@
\caption{UpSet plot for data on 132 effects from the Reproducibility Project:
  Cancer Biology \citep{Errington2021} for which effect estimates and standard
  errors are available on standardized mean difference scale. Shown are the
  number of replication successes according to the different criteria from
  Table~\ref{tab:rsdef} and their combinations. A threshold of
  $\alpha = \Sexpr{round(100*alpha, 2)}\%$ and a relative effect size threshold
  of $d_{\text{min}} = \Sexpr{dmin}$ are used for replication success (green).}
\label{fig:upset}
\end{figure}

@@

Taken together, this analysis demonstrates that conclusions based on commonly
used replication success criteria often differ. It is not always clear cut
whether or not a replication is successful. Reducing replicability to a single
criterion without mentioning these difficulties, as often done by the popular
press \citep[\eg{} ``more than half of the findings did not hold up when
retested'' in][]{Carey2015}, is a simplification of the matter.

\begin{landscape}
\begin{table}
  \centering
  \caption{Subset of results from the Reproducibility Project: Cancer Biology
    \citep{Errington2021}. (P, X, E) denotes effect number E from experiment
    number X, from original paper number P. Shown are original $\that_{o}$,
    replication $\that_{r}$, and pooled effect estimate $\that_{m}$ with 95\%
    confidence intervals, variance ratio $c = \sigma^{2}_{o}/\sigma^{2}_{r}$,
    relative effect estimate $d = \that_{r}/\that_{o}$, original
    \textit{p}-value $p_{o}$, replication \textit{p}-value $p_{r}$,
    meta-analytic \textit{p}-value $p_{m}$, $Q$-test \textit{p}-value $p_{Q}$,
    and sceptical \textit{p}-value $p_{\mathrm{S}}$. A replication success
    (green) threshold of $\alpha = 5\%$ is used for the \textit{p}-values and
    $d_{\text{min}} = 1$ for the relative effect estimate.}
  \label{tab:RPCB}
% \resizebox{1\textwidth}{!}{%
<< "RPCB-table2", results = "asis" >>=
## select a subset with one effect per paper
set.seed(7)
rpcbTable <- rpcb |>
    na.omit() |>
    ## filter(smdo < 80, smdr < 80) |>
    group_by(paper) |>
    sample_n(size = 1) |>
    ungroup() |>
    arrange(pr2) |>
    select(id,
           smdoFormat, ## so,
           smdrFormat, smdmFormat,
           c, dFormat, ## sr,
           po2Format, pr2Format, pm2Format, pQFormat, ps2Format
           )

@@

Bayesian inference is an approach to statistical inference where Bayes' theorem
is used to make probability statements about unknown parameters based on the
observed data. The central quantity for doing so is the distribution of the
parameters conditional on the data, called the \emph{posterior distribution}. It
can be obtained from Bayes' theorem,
\begin{align*}
  f(\theta \given \text{data})
  = f(\theta) \times \frac{f(\text{data} \given \theta)}{f(\text{data})},
\end{align*}
meaning that the \emph{prior distribution} for the parameter $\theta$ with
probability density/mass function $f(\theta)$ is multiplied by the (normalized)
likelihood of the data, also know as \emph{Bayesian updating}. Parameter values
which increase the likelihood of the data become more likely \emph{a posteriori}
but they are also weighted by their \emph{a priori} plausibility through the
prior. As such, Bayesian inference provides a formal way for combining
information from the data at hand with external knowledge encoded in the prior.

Many consider the prior to be also the weak point of Bayesian inference since it
is unclear how it should be specified in the absence of external knowledge. The
\emph{reverse-Bayes} approach, first proposed by \citet{Good1950}, is one way of
dealing with this issue. The idea is to reverse Bayes' theorem as
\begin{align*}
  f(\theta) = f(\theta \given \text{data}) \times
  \frac{f(\text{data})}{f(\text{data} \given \theta)}
\end{align*}
and instead ``downdate'' a posterior with the observed data. So, in contrast to
the conventional ``forward-Bayes'' approach where we start with a prior, update
it with the data, and end up with a posterior, the reverse-Bayes approach starts
with the posterior and ends up with the prior. For example, the posterior may be
specified to indicate evidence for/against a particular hypothesis under
investigation. Reverse-Bayes inference then revolves around the question of
whether the resulting prior is plausible in light of external knowledge, and if
so, this could be seen as support for the specified posterior.

To illustrate reverse-Bayes inference, let us return to the replication setting.
Assume we want to conduct inference about the unknown effect size $\theta$ based
on the effect estimate from the original study $\that_{o}$ and its standard
error $\sigma_{o}$. We will assume that $\that_{o}$ is normally distributed
around the unknown effect size $\theta$ with (known) variance equal to its
squared standard error $\sigma^{2}_{o}$, here and henceforth denoted by
$\that_{o} \given \theta \sim \Nor(\theta, \sigma^{2}_{o})$. Furthermore, we
specify a zero-mean normal prior with variance $\tau^{2}$ for the effect size
$\theta$, representing the position of a sceptic who does not believe in the
presence of a genuine effect. The ``stubbornness'' of the sceptic is determined
by how small the variance $\tau^{2}$ is chosen (the smaller $\tau^{2}$, the more
scepticism). Combining the sceptical prior with the likelihood produces a
posterior which is again normal
$\theta \given \that_{o}, \sigma_{o} \sim \Nor(\mu_{\text{post}}, \sigma^{2}_{\text{post}})$
with mean and variance
\begin{align*}
  &\mu_{\text{post}} = \frac{\that_{o}}{1 + \sigma^{2}_{o}/\tau^{2}}&
  &\text{and}&
  &\sigma^{2}_{\text{post}} = \frac{1}{1/\sigma^{2}_{o} + 1/\tau^{2}}.&
\end{align*}
The associated $(1- \alpha)$ highest posterior density credible interval is
given by
\begin{align}
  \left[\mu_{\text{post}} \pm z_{\alpha/2} \, \sigma_{\text{post}}\right]
  \label{eq:postCI}
\end{align}
and if this credible interval excludes parameter values smaller/larger than zero
(depending on the orientation of the effect size) this may be interpreted as
evidence\footnote{For readers who do not agree with this notion of evidence, do
  not panic. In this thesis (Paper II and IV) we extend this reverse-Bayes
  procedure to use alternative measures of evidence, such as Bayes factors.} for
a genuine effect found in the original study.

Depending on how large the prior variance $\tau^{2}$ is chosen, the posterior
credible interval~\eqref{eq:postCI} will either include or exclude zero.
Different researchers may have different degrees of scepticism and may thus
choose different prior variances $\tau^{2}$. As a default choice,
\citet{Held2020} proposed to use the reverse-Bayes approach from
\citet{Matthews2001b}, that is, to determine the \emph{sufficiently sceptical
  prior variance} $\tau^{2}_{\alpha}$ so that the appropriate limit of the
$(1 - \alpha)$ credible interval is just fixed to zero. The resulting prior then
represents the beliefs of a sceptic who is just stubborn enough to not find the
original study convincing at level $\alpha$.

@@

\citet{Held2019a} showed that the sufficiently sceptical prior variance
$\tau^{2}_{\alpha}$ for a level $\alpha$ is available in closed-form
\begin{align}
  \tau^{2}_{\alpha} =
  \begin{cases}
    \dfrac{\sigma^{2}_{o}}{\left(z_{o}^{2}/z_{\alpha/2}^{2}\right) - 1}
    & \text{if} ~ z_{o}^{2} > z_{\alpha/2}^{2}, \\
    \text{undefined} & \text{else.}
  \end{cases}
  \label{eq:tau2alpha}
\end{align}
From~\eqref{eq:tau2alpha} we see that convincing original studies (those with
large absolute $z$-values $|z_{o}|$) require smaller sufficiently sceptical
prior variances to render the posterior no longer convincing for the same level
$\alpha$. Conversely, if the original study is not convincing enough (if
$|z_{o}| \leq z_{\alpha/2}$) the sufficiently sceptical prior variance is
undefined meaning that the data provide so little evidence that no scepticism is
required to make them unconvincing.

@@

Figure~\ref{fig:RBillustration} shows the prior predictive $p$-values
from~\eqref{eq:pboxNorm} computed for three example replication studies from the
Replication Project: Cancer Biology. We see that larger effect estimates show
smaller prior predictive $p$-values $p_{\text{Box}}$, indicating more prior-data
conflict. This is because the standard errors from all three replications are
roughly the same size, so that the distance between zero and the replication
effect estimate matters most. The $p$-values suggest that there is hardly any
conflict between the sceptical prior and the first two replications (those with
$p_{\text{Box}} = \Sexpr{pboxFormat[1]}$ and
$p_{\text{Box}} = \Sexpr{pboxFormat[2]}$), while the conflict is larger for the
third one (the one with $p_{\text{Box}} = \Sexpr{pboxFormat[3]}$).

@@

zr <- tr/sr
c <- so^2/sr^2
ps <- pSceptical(zo = zo, zr = zr, c = c, alternative = "two.sided", type = "nominal")
psFormat <- pFormat(p = ps)
@
\subsubsection{The sceptical \textit{p}-value}
To remove the dependence on the level $\alpha$, \citet{Held2020} proposed to
determine the smallest level at which replication success can be established.
This level is called the \emph{sceptical $p$-value} $p_{\text{S}}$, and it is
available in closed-form
\begin{align*}
  p_{\text{S}}
  % = \inf_{\alpha \in (0, 1)} \left\{\alpha : p_{\text{Box}} \leq \alpha \right\}.
  = 2\left\{1 - \Phi(|z_{\text{S}}|)\right\}
\end{align*}
where
\begin{align*}
  z_{\text{S}}^{2}
  = \begin{cases}
    z^{2}_{\text{H}}/2 & \text{for} ~ c = 1 \\
    \left\{\left[z_{\text{A}}^{2}\left\{z_{\text{A}}^{2} +
          z_{\text{H}}^{2}(c - 1)\right\}\right]^{1/2}
      - z_{\text{A}}^{2}\right\}/(c - 1)
    & \text{for} ~ c \neq 1
  \end{cases}
\end{align*}
with arithmetic mean $z_{\text{A}}^{2} = (z^{2}_{o} + z^{2}_{r})/2$ and harmonic
mean $z_{\text{H}}^{2} = 2/(1/z^{2}_{o} + 1/z^{2}_{r})$ of the squared
$z$-statistics, and variance ratio $c = \sigma^{2}_{o}/\sigma^{2}_{r}$.
Replication success at level $\alpha$ is then equivalent to
$p_{\text{S}} \leq \alpha$. For instance, the sceptical $p$-values of the three
replication studies in Figure~\ref{fig:RBillustration} are
$p_{\text{S}} = \Sexpr{psFormat[1]}$, $p_{\text{S}} = \Sexpr{psFormat[2]}$, and
$p_{\text{S}} = \Sexpr{psFormat[3]}$ (from left to right), so we can see that
the third replication is unsuccessful at level $\alpha = 5\%$ but successful at
level $\alpha = 10\%$. However, the sceptical $p$-value does not necessarily
have to be dichotomized in this manner, but can also be interpreted as a
quantitative measure of replication success; the smaller $p_{\text{S}}$, the
higher the degree of replication success.

The sceptical $p$-value has several interesting properties \citep[Section
3]{Held2020}: First, it is always larger than the maximum of the original and
replication $p$-values ($p_{\text{S}} > \max\{p_{o}, p_{r}\}$), meaning that
both $p$-values have to be smaller than $\alpha$ such that replication success
at level $\alpha$ is possible. The sceptical $p$-value hence requires
\emph{both} studies to be sufficiently convincing on their own (in terms of
their $p$-values), similar to the significance criterion for replication
success. Second, if the $p$-values $p_{o}$ and $p_{r}$ remain fixed but the
relative effect estimate $d = \that_{r}/\that_{o}$ decreases, the sceptical
$p$-value increases, meaning that shrinkage of the replication effect estimate
is penalized ($p_{\text{S}} \uparrow 1$ as $d \downarrow 0$ for fixed $p_{o}$
and $p_{r}$). The sceptical $p$-value thus takes effect shrinkage into account,
unlike the significance criterion which can be achieved with any non-zero
replication effect estimate $\that_{r}$ provided the standard error $\sigma_{r}$
is small enough. This property is desirable in the replication setting as a
smaller replication effect estimate $\that_{r}$ may not be practically relevant
anymore, despite its statistical significance.

@@

This concludes the introduction to replication studies and reverse-Bayes methods
for their analysis. Some additional properties (\eg{} the null distribution of
the sceptical $p$-value) and extensions (\eg{} a one-sided version of the
sceptical $p$-value, power and sample size calculations) can be found in the
original article by \citet{Held2020}. Furthermore, the sceptical $p$-value and
related methods are implemented in the R package \texttt{ReplicationSuccess}
available on CRAN
(\url{https://cran.r-project.org/package=ReplicationSuccess}). % However, the
% relevant parts of the procedure have been presented.
In the following, I will discuss open questions and how they are addressed in
this thesis.

@@

\startPaperSummary{PaperI}{%
  The assessment of replication success based on relative effect size}{%
  Leonhard Held, Charlotte Micheloud, Samuel Pawel \\
  \textit{The Annals of Applied Statistics}, 2022, 16(2), 706--720.
  \href{https://doi.org/10.1214/21-AOAS1502}{doi:~10.1214/21-AOAS1502}.}
